Dynamics of Global and Local Social Influence

### 2021-12-01 manuscript_global_local_influence.Rmd
new revision

Individuals are adept social learners. When trying to infer the most appropriate line of action, they might use a combination of social learning strategies to extract relevant information from those around them. Our understanding of social influence, then, has to account for the fact that agents use sophisticated heuristics when they learn from one another. Our argument in this paper is straightforward: different social phenomena imply particular social learning strategies and this, in turn, affects how they spread throughout a network. Here, we focus on three strategies: relying on one's individual experience, paying attention to one's immediate ties - local influence -, and observing the most central ties of the network - global influence. We show that global and local measures are differentially informative across different behaviors. The key finding is that the emergence of distinct sub-groups, even in fully connected networks, can be explained by the weight agents place on different sources of information.

This paper has the following structure. First, we provide an overview of the theoretical issues that motivate this analysis. Second, we use the longitudinal network data-set of adolescent behavior collected by Knecht [-@knechtNetworkActorAttributes2012] to suggest that measures of global and local influence might be differentially informative across different behaviors. Third, we build an agent-based model where actors use different combinations of strategies to learn from one another. The results show that different social learning strategies lead to varying diffusion trajectories. We close by discussing the implications of these findings for the study of cultural transmission.

@@

### Influence as coordination

Peer effects are often explained by the need to “fit in” [@robalinoPeerEffectsAdolescent2018]. Individuals might adopt new habits or beliefs not only because they are exposed to novel ideas, but also because certain lines of action fit better in the social contexts where they find themselves. Imagine Jane’s school friends start taking an interest in heavy metal and going to gigs. It makes sense for her to start going to those concerts as well. It is a behavior that fits her social context: it allows her to keep spending time with her friends. This view of social influence as a “problem of fit” [@fligsteinSocialSkillTheory2001] underlies some of our most sophisticated attempts to disentangle how behaviors and ideas spread through networks. Whether it be actor-centric models of behavior dynamics [@snijdersStochasticActororientedModels2017] or spin-glass models of belief spread [@galesicStatisticalPhysicsModels2019], the core assumption is the same: actors have a higher chance of adopting habits or beliefs if such change would bring about greater synchronicity between them and the people around them. Social influence then could be conceptualized as a problem of coordination [@fligsteinSocialSkillTheory2001]. We try to orient ourselves towards the lines of action we perceive as appropriate in our social contexts.

When assessing whether actions fit an environment, we pay attention to our immediate contexts but also to more global features of the structures where we find ourselves. Suppose that Jane is part of a school where heavy metal is overall looked down upon. She can decide to only pay attention to her friends’ behavior and keep her newly acquired concert-going habits. This might result in her friendship group becoming tightly knit but socially isolated; they would be known as the ‘metalheads.’ Perhaps, she would try to balance these local and global forces by downplaying her love for loud music or complementing it by also going to pop concerts. Finding adequate lines of action in a given environment then is a balancing act between our immediate contexts and the wider structures where we are embedded.

This need to reconcile global and local forces is a central component of our widely accepted theories of social action [@abbottLinkedEcologiesStates2005; @martinWhatFieldTheory2003; @bourdieuLogicPractice1990]. Specific social settings reward certain behaviors and/or identities while discouraging others [@bourdieuLogicPractice1990]. For example, while a relentless drive to make profits might be celebrated in the corporate world, it would seem distasteful and unacceptable if embodied by an avant-garde artist. By virtue of inhabiting these social spaces, we align ourselves to their expectations and regularities. Through our piece-meal interactions, we build a picture of the patterns of behavior that configure the settings we navigate. We infer global patterns by way of our necessarily local experiences.

In the study of cultural diffusion, the interplay between local and global patterns is exemplified by ideas of “credibility” and “legitimacy” [@centolaComplexContagionsWeakness2007; @rossmanDiffusionLegitimateDiffusion2014]. The former refers to the fact that agents might abstain from adopting an innovation until they see enough of their peers embracing it. The latter, in turn, reflects that behaviors that break expectations – e.g. a radical innovation or the disadoption of a widely accepted behavior – will be met with resistance in cultural transmission. Though implicitly, these concepts highlight the fact that, when deciding whether to adopt a behavior, agents take into account what their peers are doing but also try to infer more general patterns in the population. Agents aim to infer what the information to which they have access tells them about the general acceptance of a cultural trend in the overall population. To assess the credibility or the legitimacy of an innovation is partly about figuring out how the local patterns we observe fit into the wider global dynamics.

### Beyond immediate ties: social influence and learning heuristics

Contemporary conceptions of social influence also wrestle with the tension between global and local dynamics. Increasingly, researchers have moved away from models that only consider immediate ties to more complex conceptualizations that specify varying sources of influence [@dimaggioNetworkEffectsSocial2012; @burtSocialContagionInnovation1987]. For the purposes of this analysis, it is important to highlight the work that has shown that central nodes in a structure exercise considerable social influence
76 across the network [@robalinoPeerEffectsAdolescent2018; @paluckSalienceSocialReferents2012]. These findings lend credence to the argument that popular and salient individuals are particularly influential; a view that has been furthered in scholarly work [@sherifReferenceGroupsExploration1964] but that has also been the backbone of policy interventions [@campbellInformalSchoolbasedPeerled2008]. Peer effects then are not monolithic, but rather have varying weights depending on where – and who – they come from.

The underlying assumption here is that agents are capable of inferring who these central nodes are and of drawing information from them. This is not an unwarranted hypothesis.Humans are sophisticated social learners and one of the most widely used social learning strategies, across cultures, is to copy successful individuals [@jimenezPrestigebiasedSocialLearning2019]. This means that we are highly sensitive to prestige cues. Experimental work has shown that even children engage in prestige-based social learning [@chudekPrestigebiasedCulturalLearning2012]. Salient individuals in a structure, then, are visible. They are visible because they have many connections – their in-degree count is high – but also because individuals are adept at inferring who has high status in a given social group. As Shepherd [-@shepherdStructurePerceptionHow2017] notes, these processes of inference vary depending on the social positions that individuals occupy; nodes that are highly visible to a portion of the structure might be relatively obscure to others. But we still look for prestige cues, however limited our vision might be. Agents are not passive receptacles of social influence, but they instead follow cues that might indicate where to look for useful information.

There is an established line of interdisciplinary research that examines the different strategies that individuals use when foraging for information. Imitating “successful” individuals in the network proves an important strategy when tasks are complex or when environments are highly variable [@henrichNatureCulturalTransmission2011; @mcelreathExistenceAimingOutside2008; @henrichCulturalTransmissionDiffusion2001]. When environments are relatively stable - e.g. Jane knows most people in her school will not show up tomorrow wearing Motorhead t-shirts - just looking at what the majority of group members are doing is an effective strategy [@richersonNotGenesAlone2008; @rendellWhyCopyOthers2010]. There are trade-offs between the different strategies: while identifying successful individuals entails knowledge about the markers of prestige in a given context, frequency-dependent learning, although efficient, can be ineffective in highly variable environments. Computational models and experimental data show that actors use these different strategies depending on the task at hand [@mcelreathExistenceAimingOutside2008; @barkocziSocialLearningStrategies2016]. Often the most successful strategies involve a blend of different heuristics, combined with some individual exploration.

Multiple strands of research then coalesce around the idea that agents look at different sources when trying the acquire social information. Network researchers have built increasingly complex models of social influence, while scholars interested in social learning provide evidence for the cognitive bases of these learning heuristics. What remains unclear, however, is how the use of different learning strategies affects processes of cultural transmission. For instance, recent work has used simulations to explore how beliefs, habits, and the connections between them might emerge and spread across groups. Nonetheless, these models seldom reflect the fact that agents seek information from different parts of the social structure, that they use different strategies to do so, and that they might give value to such information differently depending on the social phenomena at hand.

This line of inquiry is particularly promising because we have grounds to believe that different social learning strategies might shape how practices spread through a network. Edelmann [-@edelmannFormalizingSymbolicBoundaries2018], for instance, shows that agents respond differently to peer influence depending on how culturally proximate they consider their alters and how culturally appropriate they deem the behavior in question. In other words, if a distant acquaintance of Jane recommends a country singer, she might just dismiss it. If one of her ‘metalhead’ friends provides the same recommendation, it might carry more weight just on account of how unexpected that would be. Computational work shows that these learning biases - such as preferentially attending to in-group members [@smaldinoAdoptionSocialMarker2017] or discarding information that is too distant from one’s existent preferences [@deffuantMixingBeliefsInteracting2000] - leads to varying diffusion trajectories. Even slight changes in the information that agents take into account give way to contrasting outcomes, from total consensus to fractured polarization. The question, then, is whether foraging for information from different sources, even when no biases are present, would also lead to varying patterns of diffusion. To tackle this issue, we need two things: first, statistical methods that can help us disentangle the importance of different sources of information for a given social phenomenon; second, ways of formalizing how agents use and give weight to different social learning strategies.

@@

To measure the differential impact of different sources of influence, we can draw ideas from research on social learning, specifically on the cultural evolution of strategies in games. When a soccer manager is deciding what formation to play, they need to consider many factors: their own track-record with a strategy, how that formation has fared in the league as a whole, and the kind of strategy their rivals are likely to use [@mesoudiCulturalEvolutionFootball2020[. Recent work explores how these global and local sources of information help us predict the popularity and evolution of certain strategies across time [@mesoudiCulturalEvolutionFootball2020; @beheimStrategicSocialLearning2014]. Researchers have also designed games to conduct experimental work on broader social learning patterns [@deffnerDynamicSocialLearning2020]. Though their statistical models vary in complexity, these studies tend to use repeated measures at the individual level to fit multilevel models that examine how informative personal record and social cues are when predicting what players do next. Our argument is that we can adapt the techniques that these studies use in order to infer the informativeness of different sources of social influence in longitudinal network data.

In order to provide a formal model that specifies the importance of varying sources of influence, we can use as a starting point Friedkin and Johnsen’s [-@friedkinSocialPositionsInfluence1997] classic formulation of social influence. Formalization in this case is useful because it specifies exactly how new ideas of social influence expand upon previous work, and - importantly - the assumptions we have to make in the process. Burt's [-@burtSocialContagionInnovation1987] work remains a paradigmatic example of the usefulness of formalizing different processes of social influence. His formalization of influence through structural equivalence closely  parallels contemporary work on social learning, which shows that agents with domain expertise tend to learn from those who seem similar. In the same vein, it is useful to think how prestige-biased social learning could be incorporated into standard formal models of social influence. Friedkin and Johnsen’s [-@friedkinSocialPositionsInfluence1997] definition is as follows:

@@

What we need is an empirical strategy to adjudicate the importance of different sources of social influence for a given social phenomenon. Consider a behavior of interest like drinking alcohol. The question we would like to answer is whether to predict agent j’s drinking habits at time t + 1, we should look at the behavior of their immediate ties or at the actions of the most popular nodes of the network at time t. Using longitudinal network data, we can generate predictions using only within-person variation in global and local influence [@beheimStrategicSocialLearning2014]. Thus, this approach is useful to disentangle how informative local and global dynamics are for explaining the spread and development of a phenomena of interest – like drinking – in networks across time. In the following analysis, we fit Bayesian multilevel models to longitudinal network data to show how the informativeness of different sources of influence varies across social phenomena.

@@

Our main goal in these analyses is to examine whether these two variables – local and global scores – are differentially informative for the behaviors under examination. To address this issue, we fit Bayesian multilevel logit models where we allow the intercept and the effect of time to vary by individual. We use weakly regularizing priors that do not reflect any assumptions about the phenomena in question but rather center the estimates towards a sensible space. The prior for the effects of the global and local scores and the effect of time are narrow because overly flat priors in logit models often translate into assuming that the ends of the probability scale - either 0 or 1 - are disproportionately likely. The models we fit here have the following functional form:

>>>>

Individuals are adept social learners. When trying to infer the most appropriate line of action, they might use a combination of social learning strategies to extract relevant information from those around them. Our understanding of social influence, then, has to account for the fact that agents use sophisticated heuristics when they learn from one another. Accordingly, in this paper we offer a straightforward argument: that different social phenomena imply particular social learning strategies and this, in turn, affects how they spread throughout a network. Here we focus on three strategies: relying on one's individual experience, paying attention to one's immediate ties - local influence -, and observing the most central ties of the network - global influence. We show that global and local measures are differentially informative across different behaviors. We extend this idea using a simulation, exploring how trajectories of cultural transmission vary when agents use different social learning techniques. Our key finding is that the emergence of distinct sub-groups, even in fully connected networks, can be explained by the weight agents place on the information they retrieved using different strategies. Thus, we arrive at a result that has come to be a mark of plausible models of cultural transmission [@axelrodDisseminationCultureModel1997] by examining the interplay between social learning heuristics.

This paper has the following structure. First, we provide an overview of the theoretical issues that motivate this analysis. Second, we use the longitudinal network data-set of adolescent behavior collected by Knecht et al. [-@knechtNetworkActorAttributes2012] to suggest that measures of global and local influence might be differentially informative across different behaviors. Third, we build an agent-based model where actors use different combinations of strategies to learn from one another. The results show that different social learning strategies lead to varying diffusion trajectories. We close by discussing the implications of these findings for the study of cultural transmission.

@@

### Simulating cultural transmission and segmentation

In recent years, sociologists have used agent-based modeling to improve theories of cultural transmission between individuals. These improvements have distinguished themselves from earlier models by focusing less on the diffusion of public and material forms of culture and more on how cognitive and intrapsychic forms of culture — what Lizardo [-@lizardoImprovingCulturalAnalysis2017] has termed "personal" culture — spread through social networks. Further, this new focus has itself shifted over the last decade, as models in this vein have moved from an initial concern with explicitly articulated "declarative" beliefs [@dellapostaWhyLiberalsDrink2015] toward "nondeclarative" forms of culture, including implicit associations and social norms [@goldbergSocialContagionAssociative2018; @shawMechanicsDynamicsSocial2015; @shawSomethingOutNothing2020]. The results of this shift have been both generative and powerful, not only for explaining how actors acquire culture from others but also for how culture as shared cognitive representations [@dimaggioCultureCognition1997; @ceruloMiningIntersectionsCognitive2010] arises in the first place. Shaw [-@shawMechanicsDynamicsSocial2015], for example, presents a model in which agents in a network deploy "representations" of phenomena and observe representations deployed by their neighbors, updating their likelihood of using specific representations between iterations. As agents update which representations they use over time, group-level consensus arises around which representations should be deployed, and this consensus reinforces the use of specific representations once established. The end result is a parsimonous and empirical model of social construction in groups illustrating how individual behavior aggregates into shared culture. Moreover, Shaw finds consensus around representations in her model is rarely global: more often than not, individual representations come to prevail within specific subgroups in a network.

Since Axelrod's [-@axelrodDisseminationCultureModel1997] seminal paper, simulations of cultural transmission have sought to recreate the emergence of subgroups enacting different cultural items. This remains a  prominent feature of newer models. As Goldberg and Stein (2018:900) have argued, a common shortcoming of "contagion" models of diffusion is their implication that fully connected networks will ultimately always reach a point of full convergence, as all actors within the networks adopt and enact the same cultural item. Compared to earlier contagion models, recent simulations have prioritized the role of simple practices and heuristics in how actors process information. As individual actors use these practices and heuristics over time, they come to differ in their adoption of cultural items, culminating in segmented networks resembling those seen in the real world. Deffuant [-@deffuantMixingBeliefsInteracting2000] shows that polarization arises in contexts with no structural barriers when agents show preferences toward particular kinds of information or prefer to learn from certain types of actors. Similarly, Smaldino [-@smaldinoAdoptionSocialMarker2017] finds segmentation in networks where actors favor information coming from others with similar identities as themselves. Elsewhere, Goldberg and Stein [-@goldbergSocialContagionAssociative2018] show polarization arising when agents build associations between cultural items and use these associations to judge the same pieces of new information differently. The ability of recent simulation work to show the emergence of network polarization has vital implications for substantive work on culture and groups, as well as for broader studies of cultural diffusion. Notably, the notion that fully connected networks will always reach convergence proves false once heuristics and other practices come into account.

Recent advances in simulating cultural transmission do well to show how cultural transmission aggregates into network-wide segmentation. Nonetheless, we contend that this process can be explained using even more parsimonious models, and that doing so may be advantageous over existing alternatives. An exemplar of recent work in this area, Goldberg and Stein's [@goldbergSocialContagionAssociative2018] "associative diffusion" model features agents enacting and observing intrinsically independent pairs of cultural items. Agents form conceptual links between these pairs using a basic heuristic and modify their probability of enacting certain pairs accordingly. Over time, agents come to repeatedly enact specific pairs of items such that network-wide polarization appears. The process described here is relatively simple, and Goldberg and Stein rightfully draw from research on cognition to argue that actors form associations between concepts. At the same time, the authors assume a model of human cognition in which individual minds house networks of links between concepts, which can quickly grow complex the more cultural items they include. Research on cognition also reveals that humans are  limited in their ability to process large amounts of information. When scholars treat cultural systems as intrapsychic "webs of meaning," they risk creating networks that are too elaborate for the human mind to handle [@martinLifeBeachYou2010]. Conceptual links are undoubtedly a key part of cultural transmission, and Goldberg and Stein's model has rightfully had an important impact in this line of work. What we argue here is that it is possible to arrive at a similar finding without stepping into the cultural associations that live within agents' heads. If simpler processes produce results similar to those seen in recent work, we find ourselves with a compelling alternative to explaining how culture moves through networks and creates segmented groups.

Despite our argument that cultural transmission and polarization can be explained using simpler heuristics than recent work has offered, this work still succeeds in showing the power of relatively simple mechanisms. Yet current models of cultural transmission may be too narrow in how they conceptualize and model social learning, particularly in their tendency to deploy individual mechanisms in isolation. For the most part, these models do not incorporate multiple learning heuristics simultaneously, nor do they explore how the weights individuals place on different sources of information may shape how behaviors and ideas spread across networks. Individuals are adept social learners, and can use a combination of social learning strategies to extract relevant information from those around them when trying to infer the most appropriate line of action. Our understanding of social influence should account for the fact that actors use sophisticated combinations of simple heuristics when they learn from one another. In what follows, we argue that different social phenomena imply particular social learning strategies, and that this in turn affects how they spread throughout a network. Moreover, we argue that actors combine and weigh heuristics differently across different behaviors. Here we focus on three strategies and their combinations: relying on one's individual experience, paying attention to the behavior of one's immediate ties (local influence), and observing the behavior of the most central ties of the network (global influence).

### Individual, Local, and Global Influence in Social Settings

Research on learning trajectories in both humans and non-human animals shows the creation of preferences and habits through individual exploration. At the same time, situations arise in which learning through individual exploration is costly or intractable, prompting agents to rely on social cues for learning the best course of action. Among sociologists, this reliance on social cues is often discussed in terms of "peer effects," which are in turn explained by a need for actors to "fit into" their social environment [@robalinoPeerEffectsAdolescent2018]. Here, individuals adopt new habits or beliefs not only because of exposure to novel ideas, but also because certain lines of action fit better in the social contexts they inhabit. Imagine Jane's school friends start taking an interest in heavy metal and going to concerts. It makes sense for her to start going to those concerts as well: it is a behavior that fits her social context and allows her to keep spending time with her friends. The view of social influence as a "problem of fit" [@fligsteinSocialSkillTheory2001] underlies some of our most sophisticated attempts to disentangle how behaviors and ideas spread through networks. Whether it be actor-centric models of behavior dynamics [@snijdersStochasticActororientedModels2017] or spin-glass models of belief spread [@galesicStatisticalPhysicsModels2019], actors are assumed to have a higher chance of adopting habits or beliefs if such change would bring greater synchronicity between them and the people around them. Social influence then can be conceptualized as a problem of coordination: we try to orient ourselves towards the lines of action we perceive as appropriate in our social contexts [@fligsteinSocialSkillTheory2001].

At the same time, we also pay attention to more global features of the structures in which we find ourselves when selecting a line of action. Suppose that Jane is part of a school where heavy metal is overall looked down upon. She can still decide to only pay attention to her friends' behavior and keep her newly acquired concert-going habits. But this might result in her friendship group becoming tightly knit but socially isolated; they would be known as the 'metalheads.' Perhaps instead she would try to balance these local and global forces by downplaying her love for loud music or complementing it with other practices such as going to pop concerts. In many cases, actors strike a balance between the relative influence of their immediate contexts with that of the wider structures in which they are embedded when finding adequate lines of action.

The need to reconcile global and local forces is a central component of widely accepted theories of social action [@abbottLinkedEcologiesStates2005; @martinWhatFieldTheory2003; @bourdieuLogicPractice1990]. Specific social settings reward certain behaviors and/or identities while discouraging others [@bourdieuLogicPractice1990]. A relentless drive to make profits might be celebrated in the corporate world, but it would seem distasteful and unacceptable if embodied by an avant-garde artist. By virtue of inhabiting these social spaces, we align ourselves to their expectations and regularities. Through our piece-meal interactions, we build a picture of the patterns of behavior that configure the settings we navigate. We infer global patterns by way of our necessarily local experiences.

It remains unclear, however, how individual experience, local influence, and global influence affect processes of cultural transmission in tandem. As previously mentioned, recent simulations have tended to focus on the effects of a single mechanism in shaping the emergence and spread of beliefs, habits, and connections between them. What is more, this work has shown an almost exclusive concern with local sources of influence, with agents learning information from their immediate neighbors in a network. These models seldom reflect the fact that agents seek information from different parts of the social structure, that they use different strategies to do so, and that they might give value to such information differently depending on the social phenomena at hand. Research on transmission and the emergence of polarized networks could improve, then, from models that simultaneously account for these strategies.

An important exception to the above critique appears in DellaPosta et al.'s [-@dellapostaWhyLiberalsDrink2015] work on cultural polarization. While DellaPosta et al.'s simulation primarily looks at the effects of homophily and peer influence on polarization, the authors eventually show that the inclusion of global "hubs" of information accelerates the spread of culture through networks. Their model aims to imitate a real-world scenario of political beliefs in which highly-influential "opinion leaders" are present, and is commendable in exploring what happens when actors place different weights on multiple sources of information. Still, hubs in this model serve to represent socially distant actors: they exert influence but are not themselves influenced by other nodes in the network. We argue that global influence also occurs in more local social settings where such hubs are absent. Adolescent networks such as the one Jane finds herself in are a fitting example of this. Among adolescents, it is not only the less popular kids who adapt their behavior according to what they observe in their more popular peers: the same popular peers also look to social social cues and change their behavior based on what others are doing (Milner 2004). Accordingly, we believe that models of cultural transmission should examine sources of global influence that are more closely embedded in social settings and use the same learning strategies as other nodes in a network.

### The use of different social learning strategies

In the foregoing sections, we have argued that recent models of transmission have been limited in depicting how agents acquire and process cultural information. Specifically, these models have largely featured agents using a single heuristic to process information, and more often than not this heuristic is tied to local rather than global sources of social influence. We contend that research on cultural transmission should account for the fact that humans use multiple heuristics simultaneously when acquiring new information, including strategies that correspond to global sources of network influence. This notion finds support in an established line of interdisciplinary research that examines the different strategies individuals use when foraging for information. By drawing from this line of research, which we will refer to as the "social learning" literature, scholars of cultural transmission can modify existing models to better resemble how humans gather information while still producing similar results.

An assumption of our argument to this point has been that agents are capable of inferring which nodes in a network are most central, and that agents are able to draw information from these central nodes. Research in the social learning literature provides much evidence for this claim. Across cultures, one of the most widely used social learning strategies among humans is to copy successful individuals [@jimenezPrestigebiasedSocialLearning2019]. Humans are highly sensitive to prestige cues, and experimental work has shown prestige-based social learning occurring as early as childhood [@chudekPrestigebiasedCulturalLearning2012]. Salient and successful individuals in a structure are visible to others, in part because they have many connections (their in-degree count is high). In addition, individuals are adept at inferring high status among members composing a social group. Shepherd [-@shepherdStructurePerceptionHow2017] notes that the amount of status inferred varies depending on the social positions that individuals occupy: nodes that are highly visible to a portion of the structure might be relatively obscure to others. Nonetheless, humans still look for prestige cues within the bounds of their perception. They are not merely passive receptacles of social influence but instead follow cues based on others' positions in a group that indicate where to look for useful information.

Inferring which individuals are "successful" and imitating them is a well-documented social learning strategy based on global properties of a given social group. However, as we have suggested, this is far from the only strategy that people use when foraging for information. Social learning scholars emphasize that humans have an array of heuristics for gathering and processing information at their disposal, and that the specific heuristics humans use varies between the situations in which they find themselves. Gathering information from "successful" individuals in a network proves an important strategy when tasks are complex or when environments are highly variable [@henrichNatureCulturalTransmission2011; @mcelreathExistenceAimingOutside2008; @henrichCulturalTransmissionDiffusion2001]. For example, in a seminal ethnographic study, Henrich et. al [-@henrichNatureCulturalTransmission20] show that – among a group of Fijian villagers – perceived success in spear fishing was the biggest predictor of whether respondents would seek a person for advice regarding that activity. However, when environments are relatively stable, merely looking at what the majority of individuals in a group are doing is an effective strategy [@richersonNotGenesAlone2008; @rendellWhyCopyOthers2010]. In an experimental study, Efferson et al. [-@effersonConformistsMavericksEmpirics2008] had respondents play a game where they selected among two 'technologies' with different payoffs. A subset of participants did not have information about their individual returns but rather they knew what their immediate group had chosen - e.g. if everyone chose technology 1. For the group with social information, conformity - copying the majority - proved to be a useful strategy in the long run. There are trade-offs, of course, between different strategies. Copying successful individuals can be effective, but identifying such individuals requires knowledge about the markers of prestige in a given context. Imitating the majority, although efficient, can be ineffective in highly variable environments. Accordingly, computational models and experimental data show that actors use different strategies depending on the task and setting at hand [@mcelreathExistenceAimingOutside2008; @barkocziSocialLearningStrategies2016]. The most successful strategies often blend multiple heuristics with some individual exploration [@mcelreathExistenceAimingOutside2008]. Given these findings, cultural transmission scholars ought to consider what happens when actors blend heuristics as they often do when processing information. As different tasks and settings prompt individuals to put different weights on sources of information, they are likely to change how cultural transmission unfolds at the network level.

The social learning literature tends to focus on agents' strategies given a specific task at hand, but here we explore how the use of different learning heuristics might affect patterns of cultural transmission. The questions raised in this line of work are usually centered around the strategies themselves. In the context of simulations, researchers ask: given a complex problem, what learning strategies *should* agents use? In experimental studies, they ask: given a specific task, what heuristics do individuals *actually* use? This approach has proven useful to show that it is not only beneficial - in theory - for agents to use a varied set of learning heuristics given a problem [@rendellWhyCopyOthers2010], but that people do use them when confronted with complex tasks [@mcelreathExistenceAimingOutside2008]. Here, we try to build on this work by asking the following question: given that agents use different social learning techniques, how do these individual-level learning habits affect group-level processes of diffusion? We examine the impact that using different learning strategies might have on how culture is transmitted throughout a network. Thus, we incorporate the idea that agents use different combinations of learning strategies when foraging for information and take that as our starting point. But we also analyze how this assumption scales up to aggregate processes of cultural transmission. We examine differences in trajectories of diffusion when agents tend to rely on certain strategies or when they place different weights on the information retrieved using different heuristics.

Further, we seek to infer these learning strategies from longitudinal network data. This is an important step insofar as it grounds work on social learning in the kind of relational data that sociologists are accustomed to. If one is convinced that individuals do use different learning strategies, it follows that we would be able to see traces of them in longitudinal data. In fact, recent work on strategies used by Go players [@beheimStrategicSocialLearning2014] and soccer managers  [@mesoudiCulturalEvolutionFootball2020] has begun to look for these patterns. However, these recent efforts - as well as the simulations and experiments mentioned above - tend to only minimally incorporate social structure. Therefore, there are important benefits in testing the ideas that have emerged from the social learning literature in longitudinal network data. This type of data provides the opportunity to look at the co-evolution of behaviors and ties; to explore who is learning what and from whom, across time. Moreover, by trying to infer learning heuristics from longitudinal network data, it is possible to bridge the insights from the social learning literature with sociological work on network influence. While - as we argue below - the latter has hinted at ideas that are substantively similar as those arrived at by social learning scholars, they seldom use the same theoretical concepts. Looking for different learning strategies in longitudinal network data, then, offers a valuable extension to existing work on social learning: it grounds the insights more concretely in social structures, it helps us trace the use of strategies over time, and it extends the theoretical purchase of these ideas to other ongoing agendas.

### Different strategies among network scholars

It is somewhat surprising that most recent work on the transmission of personal beliefs and practices has not looked more explicitly at the effects of different heuristics and strategies working simultaneously. In particular, this is because network scholars elsewhere have explored how different information processing mechanisms — including those concerned with both local and global sources of peer influence — shape how other diffusion processes unfold. These scholars have increasingly moved away from models that only consider immediate ties to more complex conceptualizations that specify varying sources of influence [@dimaggioNetworkEffectsSocial2012; @burtSocialContagionInnovation1987]. Edelmann [-@edelmannFormalizingSymbolicBoundaries2018], for instance, shows that agents respond differently to peer influence depending on how culturally proximate they consider their alters and how culturally appropriate they deem the behavior in question. Closer to our argument, recent work has shown that central nodes in a structure exercise considerable social influence across the network [@robalinoPeerEffectsAdolescent2018; @paluckSalienceSocialReferents2012]. These findings lend credence to the argument that popular and salient individuals are particularly influential; a view that has both been highlighted in scholarly work [@sherifReferenceGroupsExploration1964] and has been the backbone of policy interventions [@campbellInformalSchoolbasedPeerled2008]. Peer effects then are not monolithic; rather, they have varying weights depending on where – and who – they come from.

Efforts to model multiple sources of social influence have been generative for scholars interested in how culture spreads. In the broader diffusion literature, the interplay between local and global patterns is often discussed in terms of "credibility" and "legitimacy" [@centolaComplexContagionsWeakness2007; @rossmanDiffusionLegitimateDiffusion2014]. "Credibility" here refers to the fact that agents might abstain from adopting an innovation until they see enough of their peers embracing it. "Legitimacy," in turn, reflects that behaviors that break expectations – e.g. a radical innovation or the disadoption of a widely accepted behavior – will be met with resistance in cultural transmission. Though implicitly, these concepts highlight the fact that, when deciding whether to adopt a behavior, agents take into account what their peers are doing but also try to infer more general patterns in the population. Agents aim to infer what the information to which they have access tells them about the general acceptance of a cultural trend in the overall population. To assess the credibility or the legitimacy of an innovation is partly about figuring out how the local patterns we observe fit into the wider global dynamics.

Diffusion scholars have recognized the tendency for people to use multiple heuristics or practices when acquiring and processing information, and that tensions may arise when people use heuristics and practices in combination. We argue that these tensions are valuable for models of cultural transmission that more recent work has offered and should be accounted for in this area of research. Recent work on cultural transmission has distinguished itself from broader work on diffusion with its focus on more personal and less material forms of culture. This focus has shown greater interest in - and awareness of - the cognitive bases of culture, with implications for understanding how culture moves between individuals and gives rise to group-level consensus or polarization. Accordingly, recent models of cultural transmission have operationalized the cultural items that move through networks in ways that differ from other work on diffusion. While the broader diffusion literature suggests that the interplay between different strategies and heuristics shapes the spread of personal culture, its treatment of what culture is differs enough from recent work to warrant further investigation. Consequently, we believe that newer models of cultural transmission should strive to adapt ideas explored in other research on diffusion to see how these ideas apply to beliefs, practices, and other forms of personal culture.

We also contend that this cross-pollination bears promise for network scholars. Some of the structural dynamics of social influence that this line of work outlines resonate with wider work on social learning. The influence of central nodes, for instance, can be mapped to prestige-biased social learning, while the influence of structurally equivalent alters [@burtSocialContagionInnovation1987] resembles the heuristic of copying those who are similar to us [@barkocziSocialLearningStrategies2016]. Work on diffusion takes for granted agents' social learning expertise: agents *know* who is central in a network and/or who occupies a similar position to them. What the social learning literature offers is an solid empirical basis for these assumptions. Thus, by examining different sources of social influence, through the lens of social learning, we can ground them in a wider research program that offers empirical evidence about the cognitive and cultural bases of the strategies we use to learn from one another.

In sum, network researchers have built increasingly complex models of social influence that account for individuals using multiple learning heuristics simultaneously. Outside of sociology, scholars interested in social learning provide evidence for the cognitive bases of these learning heuristics. Nonetheless, recent models of cultural transmission lag behind. These models seldom reflect the fact that agents seek information from different parts of the social structure, that they use different strategies to do so, and that they might give value to such information differently depending on the social phenomena at hand. It remains unclear then how the use of varying strategies of social learning affects processes of transmission for more personal forms of culture. To their credit, scholars of cultural transmission have found that even slight changes in the information that agents take into account give way to contrasting group-level outcomes, from total consensus to fractured polarization. The question, then, is whether foraging for information from different sources, even when no biases are present, would also lead to varying patterns in the diffusion of personal culture. This question brings us to the present study. To answer it, we need two things: first, statistical methods that can help us disentangle the importance of different sources of information for a given social phenomenon; second, ways of formalizing how agents use and give weight to different social learning strategies.

@@

To measure the differential impact of different sources of influence, we can draw ideas from research on social learning, specifically on the cultural evolution of strategies in games. When a soccer manager is deciding what formation to play, they need to consider many factors: their own track-record with a strategy, how that formation has fared in the league as a whole, and the kind of strategy their rivals are likely to use [@mesoudiCulturalEvolutionFootball2020]. Recent work explores how these global and local sources of information help us predict the popularity and evolution of certain strategies across time [@mesoudiCulturalEvolutionFootball2020; @beheimStrategicSocialLearning2014]. Researchers have also designed games to conduct experimental work on broader social learning patterns [@deffnerDynamicSocialLearning2020]. Though their statistical models vary in complexity, these studies tend to use repeated measures at the individual level to fit multilevel models that examine how informative personal record and social cues are when predicting what players do next. Our argument is that we can adapt the techniques that these studies use in order to infer the informativeness of different sources of social influence in longitudinal network data.

In order to provide a formal model that specifies the importance of varying sources of influence, we can use as a starting point Friedkin and Johnsen’s [-@friedkinSocialPositionsInfluence1997] classic formulation of social influence. Formalization in this case is useful because it specifies exactly how new ideas of social influence expand upon previous work, and - importantly - the assumptions we have to make in the process. Burt's [-@burtSocialContagionInnovation1987] work remains a paradigmatic example of the usefulness of formalizing different processes of social influence. His formalization of influence through structural equivalence closely  parallels contemporary work on social learning, which shows that agents with domain expertise tend to learn from those who seem similar [@barkocziSocialLearningStrategies2016]. In the same vein, it is useful to think how prestige-biased social learning could be incorporated into standard formal models of social influence. Friedkin and Johnsen’s [-@friedkinSocialPositionsInfluence1997] definition is as follows:

@@

What we need is an empirical strategy to adjudicate the importance of different sources of social influence for a given social phenomenon. Consider a behavior of interest like drinking alcohol. The question we would like to answer is whether to predict agent j’s drinking habits at time t + 1, we should look at the behavior of their immediate ties or at the actions of the most popular nodes of the network at time t. Using longitudinal network data, we can generate predictions using only within-person variation in global and local influence [@beheimStrategicSocialLearning2014]. Thus, this approach is useful to disentangle how informative local and global dynamics are for explaining the spread and development of a phenomena of interest – like drinking – in networks across time. In the following analysis, we fit multilevel models to longitudinal network data to show how the informativeness of different sources of influence varies across social phenomena.

@@

Our main goal in these analyses is to examine whether these two variables – local and global scores – are differentially informative for the behaviors under examination. To address this issue, we fit  multilevel logit models where we allow the intercept and the effect of time to vary by individual. We use weakly regularizing priors that do not reflect any assumptions about the phenomena in question but rather center the estimates towards a sensible space. The prior for the effects of the global and local scores and the effect of time are narrow because overly flat priors in logit models often translate into assuming that the ends of the probability scale - either 0 or 1 - are disproportionately likely. The models we fit here have the following functional form:
